---
# You need to knit from "Project Directory"
# Document information

title: "LSTAT2130 - Bayesian Statistics"
subtitle: "Project - Group Q"

authors:
  - "Lionel Lamy - 1294-1700"
  - "Adrien Kinart"
  - "Simon Lengendre"

# If multiple authors: - "Lionel Lamy - 1294-1700" is prettier.

# ---

# Logo cant have special char in the path as underscore
logo: "resources/img/UCLouvainLogoSciences.jpg"

institute: "Université catholique de Louvain"
faculty: "Louvain School of Statistics"
# department: ""

context: ""
date: \today

# ---
colorlinks: false
bordercolorlinks: true

linkcolor: "black"
urlcolor:  "black"
citecolor: "blue"

linkbordercolor: "black"
urlbordercolor: "black"
citebordercolor: "blue"

links-as-notes: false
# ---
# header-includes: 
#   -
output:
  pdf_document:
    template: template/markdown.tex
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  echo = F,
  eval = T,	
  	
  warning = F,	
  message = F,
  
  	
  out.width= "80%",
  fig.align = "center",	
  fig.path = "resources/figs/",
  
  tidy=TRUE,
  tidy.opts=list(width.cutoff=70)
)	
```

# Introduction

```{r}
library(EnvStats)
```

```{r matrix_setup}
Table <- matrix(data= c(25,69, 65,106, 80,106,136,94,76,46,
                       17,36, 47,58 , 47,53 ,59 ,54,33,21),
                nrow = 2, byrow = T)
rownames(Table) <- c("Flanders", "Wallonia") 
colnames(Table) <- c("<1200", "[1200-1500)", "1500-1800", "1800-2300", "2300-2700",
                     "2700-3300", "3300-4000", "4000-4900", "4900-6000", ">6000")
Intervals <- c(1200,1500,1800, 2300,2700,3300,4000,4900,6000)

NbFlemish <-  sum(Table[1,])
NbWaWalloons <-sum(Table[2,])  
```




```{r functions_setup}
kappaFct <- function(phi){1/phi}
lambdaFct <- function(phi,mu){1/(phi*mu)}
```

```{r flanders_wallonia}
# Flanders
Fl <- c(rep(1200,25), rep(1350,69), rep((1500+1800)/12,65), rep((1800+2300)/2,106),rep((2300+2700)/2,80), rep(3000,106),
        rep((3300+4000)/2,136), rep(4450,94),rep((4900+6000)/2,76), rep(6000,46))

Estimation_Fl <-  egamma(Fl)
Estimated_kappa_Fl <- Estimation_Fl$parameters["shape"]
Estimated_lambda_Fl <- 1/Estimation_Fl$parameters["scale"]
Estimated_mu_Fl <- Estimated_kappa_Fl/Estimated_lambda_Fl
Estimated_mu_Fl
Estimated_phi_Fl <- 1/Estimated_kappa_Fl
Estimated_phi_Fl

estimGamma_Fl <- rgamma(10000,Estimated_kappa_Fl,Estimated_lambda_Fl)
hist(Fl, probability = T, xlim = c(-500,8000))
curve(dgamma(x,Estimated_kappa_Fl,Estimated_lambda_Fl), add = TRUE)

# Wallonia
Wall <- c(rep(1200,17), rep(1350,36), rep((1500+1800)/12,47), rep((1800+2300)/2,58),rep((2300+2700)/2,47), rep(3000,53),
        rep((3300+4000)/2,59), rep(4450,54),rep((4900+6000)/2,33), rep(6000,21))


Estimation_Wall <-  egamma(Wall)
Estimated_kappa_Wal <- Estimation_Wall$parameters["shape"]
Estimated_lambda_Wal <- 1/Estimation_Wall$parameters["scale"]
Estimated_mu_Wal <- Estimated_kappa_Wal/Estimated_lambda_Wal
Estimated_mu_Wal
Estimated_phi_Wal <- 1/Estimated_kappa_Wal
Estimated_phi_Wal
estimGamma_Wal <- rgamma(10000,Estimated_kappa_Wal,Estimated_lambda_Wal)
hist(Wall, probability = T, xlim = c(-500,8000))
curve(dgamma(x,Estimated_kappa_Wal,Estimated_lambda_Wal), add = TRUE)

```



```{r JustAnExample}
muExample <- 2500; phiExample <- 1
kappaExample <- kappaFct(phiExample)
lambdaExample <- lambdaFct(muExample,phiExample)

pgamma(2700,kappaExample,lambdaExample)-pgamma(2300,kappaExample,lambdaExample)
dgamma(2700, kappaExample, lambdaExample)
```

# Question 1
Let $\theta_k:= (\mu_k, \phi_k)$ be the set of parameters for a HNI with respect to region $k$.

## (a) Theoretical probability

Let $X$ be the monthly net income of 1123 Belgian households net income (HNI) older than 30 years. Regardless the 2 regions ($k=\{1,2\}$ wrt Flanders and Wallonnia, respectively), is assumed it follows a Gamma distribution. It can be reparametrised in terms of its mean $\mu$ and dispersion parameter $\phi$ with the following trick: 


$$
\begin{split}
\text{shape: } \kappa & = \frac{1}{\phi} \\
\text{rate: } \lambda & = \frac{1}{\phi\; \mu}
\end{split}
$$


For both regions $k=\{1,2\}$: This gives


$$
f(x_k) = \frac{(\phi_k\mu_k)^{-1/\phi_k}}{\Gamma(\phi_k^{-1})} \
x_k^{1/\phi_k-1} \exp{(\frac{-x_k}{\phi_k \mu_k})}
$$

Then, the probability to fall into a certain HNI interval $I$ is: 

$$
P( x_k \in I_j) =\int_{I_j} \frac{(\phi_k \mu_k)^{-1/\phi_k}}{\Gamma(\phi_k^{-1})} x_k^{1/\phi_k-1} \exp{(\frac{-x_k}{\phi_k \mu_k})}\ \diff{x}
$$


Using CDF writing, for a region $k$ , 


\begin{equation}
p_j =
\begin{cases}
 F(x_{j}, \kappa, \lambda) \ \text{if j =1}\\
 F(x_{j+1}, \kappa, \lambda) - F(x_{j}, \kappa, \lambda) \ \text{if} \ j \in \{2,..9\}\\
 1 - F(x_{j}, \kappa, \lambda) \ \text{if j =10}\\
\end{cases}
\end{equation}



\begin{equation}
p_j =
\begin{cases}
 \frac{1}{\Gamma(\kappa)} \gamma(\kappa, \lambda x_j)  \ \text{if j =1}\\
  \frac{1}{\Gamma(\kappa)} \Big(\gamma(\kappa, \lambda x_{j+1}) - \gamma(\kappa, \lambda x_j)  \Big)   \ \text{if} j \in \{2,..9\}\\
 1- \frac{1}{\Gamma(\kappa)} \gamma(\kappa, \lambda x_j)  \ \text{if j =10}\\
\end{cases}
\end{equation}




\begin{equation}
p_j =
\begin{cases}
 \frac{1}{\Gamma(\kappa)} \int^{\lambda x_j}_0 t^{\kappa-1} e^{-t} \diff{t}   \ \text{if j =1}\\
  \frac{1}{\Gamma(\kappa)} \int^{\lambda x_{j+1}}_{ \lambda x_j} t^{\kappa-1} e^{-t} \diff{t}  \ \text{if} j \in \{2,..9\}\\
 \frac{1}{\Gamma(\kappa)} \int^{+ \infty}_{\lambda x_j} t^{\kappa-1} e^{-t} \diff{t}  \ \text{if j =10}\\
\end{cases}
\end{equation}




## (b) Theoretical expression for the likelihood
On behalf of writing simplicity, the region index is removed. Since the frequency distribution in a given region is multinomial, i.e. 

We have, writing $P:=(p{1},..,p_{10})$ and $X:= (X_{1},... X_{10})$:

$$
\begin{split}
X | P  \sim \text{Mul}(N,P)  & = \frac{x!}{x_{1}! \, ... \, x_{10}! } p_1^{x_1} \times ... \times p_{10}^{x_{10}} \; \text{when} \sum_{j=1}^{10} p_j = 1 \\
&=0 \; \text{otherwise}
\end{split}
$$

Up to a multiplicative constant, the likelihood can be written as follow:

$$
\begin{split}
L(\theta_k, D_k) = P (D_k | \mu_k, \phi_k) \propto \prod_{j=1}^{10} p_{k,j}^{x_{k,j}}
\end{split}
$$

## Taking approximation 

$p_j$ corresponds to the area in the $j^{\text{th}}$ interval. One can take the approximation mean the mean,e.g. $x_{Flanders,3} = (1500+1800)/2 = 1650$ ? On can approximate that with $f(x_i) \Delta$ where $\Delta$ is the unit of measurement. 

$$
\begin{split}
p_j =P(x_j - \frac{\Delta_j}{2} < x_j <x_j + \frac{\Delta_j}{2} ) & \approx f(x_j) \Delta_j \\
& \approx \frac{1}{\Gamma(\phi_k^{-1})} \big( \phi_k \mu \big)^{-1/\phi_k} x_j^{\frac{1}{\phi_k}-1} \exp{(\frac{-x_j}{\phi_k \mu})} \Delta_j
\end{split}
$$
This gives for the likelihood: 


$$
\begin{split}
P (D | \mu, \phi) & \propto \prod_{j=1}^{10}  x_j^{\frac{1}{\phi_k}-1} \exp{(\frac{-x_j}{\phi_k \mu})} \Delta_j \\
& \propto \exp{(\frac{- \sum x_j}{\phi_k \mu})}  \prod_{j=1}^{10}  x_j^{\frac{1}{\phi_k}-1}  \Delta_j \\
\end{split}
$$

## Not taking approximation but the CDF differences


$$
\begin{split}
P(D| \kappa, \lambda) & \propto \gamma(x_1 , \kappa, \lambda)^{x_1} \Bigg[ \prod_{j=2}^9  \Big( \gamma(x_{j} , \kappa, \lambda)- \gamma(x_{j-1} , \kappa, \lambda)\Big)^{x_j}\Bigg] \big(1-\gamma(x_{10} , \kappa, \lambda)\big)^{x_{10}}  \\
& \propto \Bigg( \int_0^{\lambda x_1}  x^{\kappa-1} e^{-x} \diff x \Bigg)^{x_1}  \Bigg[ \prod_{j=2}^9 \Big( \int_{\lambda  x_j-1}^{\lambda x_{j}}  x^{\kappa-1} e^{-x} \diff x \Big)^{x_{j}}\Bigg] \Bigg( \int_{\lambda x_{10}}^{+\infty}  x^{\kappa-1} e^{-x} \diff x \Bigg)^{x_{10}}
\end{split} 
$$

If we write $[0; x_1], [x_1; x_2], ..., [x_9; x_{10}], [x_{10}, +\infty]$ as $I_1, I_2, ..., I_9$ and $I_{10}$. The notation can be lightened. With respect to the region $k$, this gives:
$$
P(D_k | \kappa_k, \lambda_k) \propto \prod_{j=1}^{10} \Bigg( \int\limits_{\lambda_k I_j} x^{\kappa_k-1} e^{-x} \diff x \Bigg)^{x_{k,j}}
$$

Writing in terms of $\mu_k$ and $\phi_k$: 

$$
P(D_k | \mu_k, \phi_k) \propto \prod_{j=1}^{10} \Bigg( \int\limits_{ \frac{I_j}{\phi_k \mu_k}} x^{(\phi_k^{-1}-1)} e^{-x} \diff x \Bigg)^{x_{k,j}}
$$


# Question 2 : Priors 

- Statement 1: we are at 95% convinced that the mean net monthly household income in a given region is in the interval (2400, 3600). If one assumes a normal distribution for the mean $\mu_k$ (à justifier), then it is possible to get a prior of the distriubtion for both regions: 

For both regions $\mu_0 = 3000$. Then, to get the standard deviation: 

$$
\begin{split}
& 3000 - t_{(n_k-1, 1-\alpha/2)} \frac{s_k}{\sqrt{n_k}} = 2400 \\
\rightarrow  & \hat{\sigma}_0 = \frac{s_k}{\sqrt{n_k}}= \frac{600}{t_{(n_k-1, 1-\alpha/2)}}
\end{split}
$$


$\hat{\sigma}_{Fl} = 306$

So we have 
$$\mu \sim N(\mu_0= 3000, \sigma_0 =306)$$

$$
\begin{split}
 \pi(\mu) \propto \sigma_0^{-1/2} \exp{ \big(-\frac{1}{2\sigma_0^2}(\mu-\mu_0)^2\big) }
\end{split}
$$


- dispersion parameter: 

If one considers that the parameter can be in any point within the interval (0.0, 10.0) with the same probability, then one could say that it follows a uniform distribution between those to interval 

$$
\phi_k \sim U (a=0,b=10) \propto 1_{0,10}
$$

Using the entropy theorem, the conjugate prior would be the product of the two last quantity: 


$$
\begin{split} 
\text{prior:}  P(\mu_k, \phi_k) & = P(\mu_k) \ P(\phi_k) \\
& \propto \sigma_0^{-1/2} \exp{ \big(-\frac{1}{2\sigma_0^2}(\mu-\mu_0)^2\big) } 1_{0;10}
\end{split}
$$

# Question 3

## Question 3a: posterior

### With approximation 
$\sigma^2 = \phi \mu^2$

$$
\begin{split}
P(\mu, \phi | D) & \propto \exp{(\frac{- \sum x_j}{\phi_k \mu})}  \prod_{j=1}^{10}  x_j^{\frac{1}{\phi_k}-1}  \Delta_j \sigma^{-1/2} \exp{ \big(-\frac{1}{2\sigma^2}(\bar{x}_k-\mu)^2\big) } 1_{0,10} \\ 
& \propto \exp{(\frac{- \sum x_j}{\phi_k \mu})}  \prod_{j=1}^{10}  x_j^{\frac{1}{\phi_k}-1}  \Delta_j (\frac{1}{\sqrt{\phi} \mu}) \exp{ \big(-\frac{1}{2 \phi \mu^2}(\bar{x}_k-\mu)^2\big) } 1_{0,10} \\
\end{split}
$$


### Without approximation

$$
\begin{split}
P(\mu_k, \phi_k | D) \propto \Bigg( \prod_{j=1}^{10} \Bigg( \int\limits_{ \frac{I_j}{\phi_k \mu_k}} x^{(\phi_k^{-1}-1)} e^{-x} \diff x \Bigg)^{x_{k,j}} \Bigg)  \sigma_0^{-1/2} \exp{ \big(-\frac{1}{2\sigma_0^2}(\mu-\mu_0)^2\big) } 1_{0;10}
\end{split}
$$

## 3.b 

the log likelihood is:
$$
\begin{split}
LL & \propto \sum_{j=1}^{10} x_j \ln{\Bigg( \int\limits_{ \frac{I_j}{\phi_k \mu_k}} x^{(\phi_k^{-1}-1)} e^{-x} \diff x \Bigg)}
\end{split}
$$

```{r}
lpost <- function(theta,freq){
  Intervals <- c(1200,1500,1800, 2300,2700,3300,4000,4900,6000) # length =9 
  mu <- theta[1]
  phi <- theta[2]
  kappa <- 1/phi
  lambda <- 1/(phi*mu)
  n <- length(Intervals)

  #initialisation: j=1 ==>  x_1 * proba(0,x_1) 
  LL <- freq[1]* log(pgamma(Intervals[1], kappa, lambda))
  
  for (i in 2:(n-1)) {
    LL <- LL + freq[i] *log(pgamma(Intervals[i+1], kappa, lambda) - pgamma(Intervals[i], kappa, lambda))
  }
  ## de x_n  à +infinity (=> 6000 --> inf )
  LL <-  LL+ freq[n]* log(1-pgamma(Intervals[n], kappa, lambda))
  return(LL)
}
```

```{r}
lpost(c(Estimated_mu_Fl, Estimated_phi_Fl), t(Table[1,]))
lpost(c(Estimated_mu_Wal, Estimated_phi_Wal), t(Table[2,]))
```


# 4. 


$$
h(\mu,\phi) \propto  l(\mu,\phi) + \ln(\pi(\mu,\phi))
$$

$$
h(\mu,\phi) = Ct + \sum_j x_j \ln{\big(F(\mu,\phi,x_{j+1})-F(\mu,\phi,x_{j})  \big)} - \frac{1}{2 \sigma_0^2} (\mu-\mu_0)^2
$$

... 
pour le moment... (voir photo envoyée sur messenger)
$$
\begin{split}
\mu^* & =  \mu_0 + \sigma_0^2 (\sum^{10}x_j \frac{f_{j+1} - f_{j}}{ F_{j+1} -F{j}} ) \\
\sigma^{2*} &= \sigma_0^2 
\end{split}
$$

<!------>
\appendix

# Appendix{#appendix}
## Figures{#figures}
## Code{#code}

\bigskip
\begin{mdframed}[style=thicc, frametitle=Note, frametitlebackgroundcolor=black!30]
  For reproducibility purposes, the complete R project containing the source code and the results is available on \href{https://github.com/AdrienKinart/LSTAT2130BayesianProject}{github.com}.
\end{mdframed}


```{r appendix_pre}
# remotes::install_github('yihui/formatR')
library(formatR)
```

```{r appendix_code, eval=FALSE, echo=TRUE, ref.label=knitr::all_labels(appendix==T)}
```
