---
# Document information

title: "LSTAT2130 - Bayesian Statistics"
subtitle: "Project - Group Q"

authors:
  - "Lionel Lamy - 1294-1700"
  - "Adrien Kinart"
  - "Simon Lengendre"

# If multiple authors: - "Lionel Lamy - 1294-1700" is prettier.

# ---
# You can remove settings below
# Defaults values are defined

# logo: ""

institute: "Université catholique de Louvain"
faculty: "Louvain School of Statistics"
# department: ""

context: ""
date: \today

# ---
colorlinks: false
bordercolorlinks: true

linkcolor: "black"
citecolor: "blue"
urlcolor:  "blue"

linkbordercolor: "black"
urlbordercolor: "blue"
citebordercolor: "blue"

links-as-notes: false
# ---
# header-includes: 
#   -
output:
  pdf_document:
    template: template/markdown.tex
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  echo = F,
  eval = T,	
  	
  warning = F,	
  message = F,
  
  	
  out.width= "80%",
  fig.align = "center",	
  fig.path = "resources/figs/",
  
  tidy=TRUE,
  tidy.opts=list(width.cutoff=70)
)	
```

# Introduction

```{r}
Table <- matrix(data= c(25,69, 65,106, 80,106,136,94,76,46,
                       17,36, 47,58 , 47,53 ,59 ,54,33,21),
                nrow = 2, byrow = T)
rownames(Table) <- c("Flanders", "Wallonia") 
colnames(Table) <- c("<1200", "[1200-1500)", "1500-1800", "1800-2300", "2300-2700",
                     "2700-3300", "3300-4000", "4000-4900", "4900-6000", ">6000")

NbFlemish <-  sum(Table[1,])
NbWaWalloons <-sum(Table[2,])  
```




```{r}
kappaFct <- function(phi){1/phi}
lambdaFct <- function(phi,mu){1/(phi*mu)}
```



```{r JustAnExample, echo=FALSE}
muExample <- 2500; phiExample <- 1
kappaExample <- kappaFct(phiExample)
lambdaExample <- lambdaFct(muExample,phiExample)

pgamma(2700,kappaExample,lambdaExample)-pgamma(2300,kappaExample,lambdaExample)
```

# Question 1
Let $\theta_k:= (\mu_k, \phi_k)$ be the set of parameters for a HNI with respect to region $k$.

## (a) Theoretical probability

Let $X$ be the monthly net income of 1123 Belgian households net income (HNI) older than 30 years. Regardless the 2 regions ($k=\{1,2\}$ wrt Flanders and Wallonnia, respectively), is assumed it follows a Gamma distribution. It can be reparametrised in terms of its mean $\mu$ and dispersion parameter $\phi$ with the following trick: 


$$
\begin{split}
\kappa & = \frac{1}{\phi} \\
\lambda & = \frac{1}{\phi\; \mu}
\end{split}
$$


For both regions $k=\{1,2\}$: This gives


$$
\begin{split}
f(x_k) = \frac{1}{\Gamma(\phi_k^{-1})} \big( \phi_k \mu \big)^{\phi_k} x_k^{\frac{1}{\phi_k}-1} \exp{(\frac{-x_k}{\phi_k \mu})}
\end{split}
$$


Then, the probability to fall into a certain HNI interval is: 

$$
\begin{split}
P(x_{j_1} < x < x_{j_2}) =\int_{x_{j_1}}^{x_{j_2}} \frac{1}{\Gamma(\phi_k^{-1})} \big( \phi_k \mu \big)^{\phi_k} x_k^{\frac{1}{\phi_k}-1} \exp{(\frac{-x_k}{\phi_k \mu})}
\end{split}
$$



## (b) Theoretical expression for the likelihood

We have, writing $P:=(p_{k,1},..,p_{k,10})$ and $X_k:= (x_{k,1},... x_{k,10})$
$$
\begin{split}
X_k |P  \sim \text{Mul}(n_k,P)  & = \frac{x_k!}{x_{k,1}! \, ... \, x_{k,10}! } p_1^{n_{x,1}} \times ... \times p_1^{x_{k,10}} \; \text{when} \sum_{j=1}^{10} x_j = x_k \\
&=0 \; \text{otherwise}
\end{split}
$$

$$
\begin{split}
L(\theta_k, D_k) = P (D_k | \theta_k) = \pi_{}
\end{split}
$$
<!------>
\appendix

# Appendix{#appendix}
## Figures{#figures}
## Code{#code}

\bigskip
\begin{mdframed}[style=thicc, frametitle=Note, frametitlebackgroundcolor=black!30]
  For reproducibility purposes, the complete R project containing the source code and the results is available on https://github.com/.
\end{mdframed}


```{r appendix_pre}
# remotes::install_github('yihui/formatR')
library(formatR)
```

```{r appendix_code, eval=FALSE, echo=TRUE, ref.label=knitr::all_labels(appendix==T)}
```
